---
title: "Final Project RMD"
author: "Marali Benitez, Youcef Djoudi, Kelly Thackery, Samina Issa, Samira Bechi"
date: "2025-04-09"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
##loading in packages and data
library(wooldridge)
library(AER)
library(quantmod)

#ceosal2 data
ceosalary <- ceosal2

#CPS 1988 data
data("CPS1988")
cps <- CPS1988
```

**Data: ceosal2**

**(a) Describe the data. Describe the variables; response, predictors, continuous, categorical variables and missing data**.   



**(b) Build an optimal model. Print summary: discuss F-test, t-tests, Rsquare, S-square. Take out outliers**.  



**(c) Briefly discuss residuals**.  



**(d) Make 2 predictions with CI's. Interpret predictions if needed**.  




**Data: CPS 1988**.   

**(a) Describe the data. Describe the variables; response, predictors, continuous, categorical variables and missing data**.   



**(b) Build an optimal model. Print summary: discuss F-test, t-tests, Rsquare, S-square. Take out outliers**.  



**(c) Briefly discuss residuals**.  



**(d) Make 2 predictions with CI's. Interpret predictions if needed**.  